Using cosmic variance to constrain the dark matter halo mass and occupation fraction of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-06-26 paper/ms.tex
in better shape for the discussion

From the matching of number density statistics we find that all
the succesful models have halos in the range
$10^{10}-10^{12}$. Including strict criteria and the information from
the angular correlation function does not narrow this mass range, but
help us nevertheless to narrow down the number of good models.

matter masses of halos hosting LAEs is $\log_{10}M_{\rm
  med}=10.9^{+0.5}_{-0.9}$\Msun, with a corresponding occupation
fraction of $1-10\%$.  Our results are in a good agrement with those estimates for the
host mass. This is not unexepected given that we have also
required consistency with ACF measurements.

Together with the escape fraction $f_{\rm occ}$, quantifying the halo mass range
$\Delta M=\log_{10}M_{\rm max} - \log M_{\rm  min}$ for the succesful
models help us to classify them into three families:

@@

There is a clear majority of models with a narrow $\Delta M\leq
1$. Such models imply that there is a cut for lower and higher halo
masses that render inefficient the presence and/or detection of LAEs.
A cut at the low mass end can be readily interepreted in terms of the
minimal halo star formation rate needed to produce the necessary \lya
luminosity to be above a given detection threshold.  However, under the reasonable assumption of star formation rate
increainsng with halo mass, the cut at higher halo masses requires a
different explanation.

A possible interpretation can be made in terms of a drecreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents than lower mass
systems, which due to the resonant nature of the \ly line are enough
to produce high absorption of \ly photons but not of continuum or
other non-resonant lines.

This preference for narrow $\Delta M$ ranges for hosting LAEs together
with the total absence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun shows that our models supports theoretical insights
where the most massive systems are not bright \ly sources \citep{ForeroRomero2012}.

We find that $70\%$ of the best models have a low
occupation fraction $f_{\rm occ}\leq 0.3$.

With all the observational constraints we we end up with 50
models out of the initial 90000 combination of parameters. In what
follows we classify the resulting models into three families and
discuss them under the light of previous observational analysis from
blind galaxy surveys and other LAEs clustering analysis.

\subsection{Constraints from blind surveys}

This preference goes in
the same direction as the observational constraint on $f_{\rm occ}\sim
0.1-0.2$ derived at $z=2.2$ by \cite{Hayes2010} and recent theoretical
study of observational data in a wide redshift range $0<z<6$ 
\citep{Dijkstra2013}.

@@

\cite{Dijkstra2013} use a similar principle to derive their results. They
compared observationally derived star formation functions to LAE
luminosity functions. At $z\sim 3.0$ they derive an affective escape
fraction of $f_{\rm esc}=(17\pm 5)\%$ could be interpreted as an
occupation fraction $f_{\rm occ}\sim 0.2$.  We consider a sucess of
our method the fact that we find that most of the consistent models
show a low occupation fraction.

@@

Since the observational estimates based on the ACF infered from photometric
measurements in the Extended Chandra Deep Field South
\citep{Gawiser07} it is a well accepted fact that the median dark
matter masses of halos hosting LAEs is $\log_{10}M_{\rm
  med}=10.9^{+0.5}_{-0.9}$\Msun, with a corresponding occupation
fraction of $1-10\%$.  Our results are in a good agrement with those estimates for the
host mass. This is not unexepected given that we have also
required consistency with ACF measurements.

The novelty in our approach is that we have a detailed estimate for 
host halo mass range. Following the numbers shown for the three
families in the previous section, we see that two thirds of the best models
have a narrow mass range, less than $<1.0$ dex, out of the $2.5$ dex
spanned by dark matter halos at that redshift. Only one fourth of the
models indicate a wide mass range with $>1.0$dex.

As a byproduct of our results, we are now able to demonstrate that a
simple clustering analysis is not enough to rule out models with a
high occupation fraction $f_{\rm occ}>0.3$, they represent almost
one fourth of our best models, coinciding with a wide range in halo
masses $\Delta M>1.0$ dex.

%\subsection{In the context of abundance matching models}

%Using abundance matching methods throughout cosmic time from redshifts
%$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} report that the instantaneous star
%formation efficiency (star formation rate divided by the stellar mass)
%presents a clear maximum around $10^{11.7}$\Msun at all redshifts
%$z<4$.

%This mass scale is strictly superior to the all the $M_{\rm max}$
%values allowed in the consistent models. Using the results published
%in \cite{Behroozi2013a} we find that the typical stellar mass in halos
%of $10^{11.4}$\hMsun is $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while
%their star formation rate is in the range $0.6\pm 0.2$ Msun
%yr$^{-1}$.  This halo mass range around $10^{11.4}$ \hMsun is the
%lower bound of what is computed in the abundance matching model in
%order to fit the observational data for the stellar mass function
%based on the obervations of Lyman Break Galaxies.

%Most of the prefered models have a halo mass range with $M_{\rm
%  max}<10^{11.4}$\hMsun. This suggests that a detailed and careful
%study of the spectral properties of LAEs coupled to the kind of
%analysis performed in this paper can be a guide in the study of the
%properties of low mass dark matter halos at $z=3.1$, extending the
%capabilities of abundance matching methods.

%we have a wide range in halo masses, where the upper end of the halo
%masses can be considered as observed LAEs, one can expect that bright
%LBGs will have a correspondence as observed LAEs. The most interesting
%aspect is that there is a clear cut in the minimal mass that can be
%attained by observed LAEs $M_{\rm min}> 10^{11}$\hMsun. This puts a tight
%constraint on the relationship between the minimum star formation rate
%required to be observed as a LAE and this minimal halo mass.

>>>>

Matching the galaxy surface number density statistics sets the median
mass of all succesful models in the range $10^{10}-10^{12}$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 90000 possible combination of parameters.

We find that for the physical interpretation of these models it is
useful to do it in terms of the size of halo mass range, using the
variable $\Delta M=\log_{10}M_{\rm max} - \log M_{\rm  min}$, together
with the escape fraction $f_{\rm occ}$. These two variables help us
build a classification of all the successful models into three families:

@@

There is a clear majority of models with a narrow $\Delta M\leq
1$, compared to the $2.5$dex of halo mass available for occupation at
that redshift. Such models imply that there is a cut at lower and higher halo
masses that render inefficient the presence and/or detection of LAEs.

At the low mass end, such cut can be readily interepreted in terms of the
minimal halo star formation rate needed to produce the necessary \lya
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increainsng with halo
mass, the cut at higher halo masses requires a different explanation.

One possible interpretation can be made in terms of a drecreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents than lower mass
systems, which due to the resonant nature of the \ly line are enough
to produce high absorption of \ly photons but not of continuum or
other non-resonant lines.

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the total absence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun shows that our models supports theoretical insights
where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}.

\subsection{Comparison against results from blind surveys}

We also find that $70\%$ of the best models are found in families (1)
and (2), with a low occupation fraction $f_{\rm occ}\leq 0.3$. This
preference goes in the same direction as the observational constraint
on $f_{\rm occ}\sim 0.1-0.2$ derived at $z=2.2$ by \cite{Hayes2010}
and recent theoretical study of observational data in a wide redshift
range $0<z<6$  \citep{Dijkstra2013}.

@@

\cite{Dijkstra2013} used a similar principle to derive their results. They
compared observationally derived star formation functions to LAE
luminosity functions. At $z\sim 3.0$ they derive an affective escape
fraction of $f_{\rm esc}=(17\pm 5)\%$ could be interpreted as an
occupation fraction $f_{\rm occ}\sim 0.2$.  We consider a sucess of
our method the fact that we find that most of the consistent models
show a low occupation fraction.

@@

Observational based on the ACF infered from photometric measurements
in the Extended Chandra Deep Field South have shown that the median
dark matter masses of h los hosting LAEs is $\log_{10}M_{\rm
  med}=10.9^{+0.5}_{-0.9}$\Msun, with a corresponding occupation
fraction of $1-10\%$  \citep{Gawiser07}.  Our results are in a general
good agrement with those estimates for the host mass. This is not
completeley unexepected given that we have also required consistency
with ACF measurements.

The novelty in our approach is that we have a detailed estimate for 
host halo mass range together with the escape fraction. This allows us
to demonstrate that the halo mass range could be very narrow $\Delta M <
0.2$dex, something that cannot be inferred from ACF analysis
alone.

WE also find interesting that ACF analsys is also not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models can only be
considered unfavorable based on a different set of observations as we
have described in the previous subsection.

\subsection{In the context of abundance matching models}

Considering the additional evidence for a low escape fraction we can
say that the prefered models are in families (1) and (2) with a
clear majority composed by those with narrow mass range and low occupation
fraction.

Using abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} report that the instantaneous star
formation efficiency (star formation rate divided by the stellar mass)
presents a clear maximum around $10^{11.7}$\Msun at all redshifts
$z<4$.

This mass scale is strictly superior to the great majority of  $M_{\rm max}$
values allowed in the models with low escape fraction. Using the results published
in \cite{Behroozi2013a} we find that the typical stellar mass in halos
of $10^{11.4}$\hMsun is $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while
their star formation rate is in the range $0.6\pm 0.2$ Msun yr$^{-1}$.
This halo mass range around $10^{11.4}$ \hMsun is the lower bound of
what is computed in the abundance matching model in order to fit the
observational data for the stellar mass function based on the
obervations of Lyman Break Galaxies.

All of the prefered models have a halo mass range with $M_{\rm
  min}<10^{11.4}$\hMsun. This suggests that a detailed and careful
study of the spectral properties of LAEs coupled to the kind of
analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.
